Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-03-03 Manuscript/HI_Model_ms.Rmd
updated manuscript files

### *4.1 Impact of different climate and land-use pathways* 
By comparing ecosystem carbon balance estimates under different scenario combinations, we were able to assess the relative impact of both global emissions reductions and regional actions to reduce emissions from land use, land use change, and disturbances such as wildland fire (figure 5). Global adherence to a lower emissions trajectory (i.e., switching from RCP 8.5 to RCP 4.5) had the largest cumulative impact, resulting in a median of 26 additional Tg C sequestered by Hawaiian ecosystems over the 90-year simulation period. Long-term reductions in the intensity of land-use change also consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global emissions reductions. Switching from the ‘high’ to the ‘low’ land‐use scenario resulted in a median cumulative retention of an additional 11.6 Tg C in Hawaiian ecosystems by 2100. The combination of global climate mitigation and a reduction in land‐use conversions had the largest potential benefit to ecosystem carbon sequestration, reducing cumulative net losses by over 400% (37.7 Tg C). Notably, the relative impact of reducing emissions from land-use change was much greater under the high radiative forcing pathway (RCP 8.5). Cumulative NBP increased by 130% when switching from the high to low land-use scenario under RCP 8.5, as opposed to a 37% cumulative increase in NBP when switching from high to low land-use under RCP 4.5.

### *4.2 Comparison to other studies*
There are few estimates of contemporary ecosystem carbon balance for the main Hawaiian Islands, and even fewer model projections of future ecosystem carbon balance in response to climate and land use change.  Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2020 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019]. In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period {0.314 Tg C y^-1^; @selmants_baseline_2017]. Our current model estimates of NPP and R~h~ being 9.5% and 15% lower, respectively [

>>>>

### *4.1 Impact of different climate and land-use pathways* 
By comparing ecosystem carbon balance estimates under different scenario combinations, we were able to assess the relative impact of both global emissions reductions and regional actions to reduce emissions from land use, land use change, and wildland fire (figure 5). Global adherence to a lower emissions trajectory (i.e., switching from RCP 8.5 to RCP 4.5) had the largest impact, resulting in a median cumulative increase of 26 Tg C sequestered by Hawaiian ecosystems over the 90-year simulation period. Long-term reductions in the intensity of land-use change also consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global emissions reductions. Switching from the high to the low land use scenario resulted in a median cumulative retention of an additional 11.6 Tg C in Hawaiian ecosystems by 2100. The combination of global climate mitigation and local reductions in land use conversion had the largest potential benefit to ecosystem carbon sequestration, reducing cumulative net losses by over 400% (37.7 Tg C). Notably, the relative impact of reducing emissions from land-use change was much greater under the high radiative forcing pathway (RCP 8.5). Cumulative NBP increased by 130% when switching from the high to low land-use scenario under RCP 8.5, as opposed to a 37% cumulative increase in NBP when switching from high to low land-use under RCP 4.5. These results demonstrate that reducing ecosystem carbon losses from land use change, harvest, and wildland fire can be an important component of greenhouse gas reduction efforts by sub-national jurisdictions like the State of Hawaiʻi, regardless of the global emissions trajectory. These results also highlight the utility of Hawaiʻi's multi-pronged approach of participating in global climate mitigation efforts by reducing emissions from the energy and transportation sectors while also reducing land-use emissions to minimize positive feedbacks to the climate system.

### *4.2 Comparison to other studies*
There are few estimates of contemporary ecosystem carbon balance for the main Hawaiian Islands, and even fewer model projections of future ecosystem carbon balance in response to climate and land use change.  Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2020 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019]. In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.314 Tg C y^-1^; @selmants_baseline_2017], despite our current model estimates of contemporary NPP and R~h~ being 9.5% and 15% lower, respectively.
